---
title: "p8105_hw3_jmm2511"
author: "Jessie Merrifield"
date: "2023-10-14"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 7,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 2


format the data to use appropriate variable names;

focus on the “Overall Health” topic

include only responses from “Excellent” to “Poor”

organize responses as a factor taking levels ordered from “Poor” to “Excellent”



```{r Q2 read and clean}
data("brfss_smart2010")

brfss_smart2010<-brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic=="Overall Health", 
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  mutate(response=factor(response, ordered = TRUE, levels=c("Poor","Fair", "Good","Very good","Excellent")))

```

In 2002, which states were observed at 7 or more locations? What about in 2010?

In 2002, CT, FL, MA, NC, NJ, and PA were observed at 7 or more locations.
In 2010, CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA were observed at 7 or more locations

```{r}
brfss_smart2010 %>% 
 filter(year=="2002") %>% 
 group_by(locationabbr) %>% 
 summarise(n_location=n_distinct(locationdesc)) %>% 
  filter(n_location>=7)

brfss_smart2010 %>% 
 filter(year=="2010") %>% 
 group_by(locationabbr) %>% 
 summarise(n_location=n_distinct(locationdesc)) %>% 
  filter(n_location>=7)

```

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 

```{r Q2 excellent df}

brfss_excellent<-brfss_smart2010 %>% 
  filter(response=="Excellent") %>% 
  group_by(year, locationabbr) %>% 
  summarise(mean_data_value=mean(data_value))

```

Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r spaghetti plot}
brfss_spaghetti<-
  brfss_excellent %>% 
  ggplot(aes(x=year, y=mean_data_value,color=locationabbr))+
  geom_line()+ 
  labs(
    title = "Average data values from 2002 to 2010 by state",
    x = "Year",
    y = "Average data value",
    color = "State")

brfss_spaghetti
```


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State

```{r NY plot}

brfss_NY_plot<-
  brfss_smart2010 %>% 
  filter(locationabbr=="NY", year== c("2006", "2010")) %>% 
  ggplot(
    aes(x=data_value))+
  geom_density(aes(fill=response), alpha=0.5)+
  facet_grid(~year)+ 
  labs(
    title = "Data value density for responses from 2006 and 2010",
    x = "Data value",
    y = "Density",
    color = "Response")

brfss_NY_plot

```


### Problem 3

Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind)

```{r Q3 read merge}
nhanes_accel<-read.csv("./data/nhanes_accel.csv") %>% 
  janitor::clean_names() 

nhanes_covar<-read.csv("./data/nhanes_covar.csv",
                       skip = 4) %>% 
  janitor::clean_names() %>% 
  filter(age>=21) %>% 
  drop_na() %>% 
  mutate(
    sex=recode(sex, "1"="male", "2"="female"),
    education=recode(education, "1"="Less than high school", "2"="High school equivalent",
                     "3"="More than high school")
  )%>% 
  mutate(
    sex=factor(sex, ordered = TRUE, levels=c("male","female")),
    education=factor(education, ordered = TRUE, c("Less than high school","High school equivalent","More than high school")))

#before joining--excluded participants with less than 21 years of age, those with missing data from nhanes_covar (demographic), and changed sex and education into ordered factors. 

#used left_join to keep the seqn present in the nhanes_covar dataset with 228 obs and a final 1445 variables

#used pivot longer to tidy the data 

nhanes_joined<-
  left_join(nhanes_covar, nhanes_accel, by="seqn")


nhanes_long<-nhanes_joined%>% 
    pivot_longer(
    min1:min1440,
    names_to = "minute",
    names_prefix = "min",
    values_to = "MIM_values"
  )


```

Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items

```{r sex and education distribution}
nhanes_joined %>% 
  group_by(education, sex) %>% 
  summarise(n=n()) %>% 
  pivot_wider(
    names_from = sex,
    values_from = n
  ) %>% 
  knitr::kable(digits = 1)
# the number of females in each category is similar to males except for there being a greater number of males in the "high school equivalent" category

age_distribution<-
  nhanes_joined %>% 
  ggplot(
    aes(y=age, x=education, color=sex))+
  geom_boxplot()+ 
  labs(
    title = "Age distribution by education level and sex",
    x = "Education",
    y = "Age",
    color = "Sex")

age_distribution


#The average ages of males decreases with increasing education leves. For females, the highest ages are in the high school equivalent level, followed by the less than high school education level, and finally the more than high school education level. For both sexes, the lowest ages are within the more than high school education level. 

```

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

```{r}


total_activity<-
  nhanes_joined %>%
  group_by(seqn, sex, age, education) %>% 
  summarise(total_activity=sum(min1:min1440)) %>% 
  ggplot(aes(x=age, y=total_activity, color=sex))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  facet_grid(.~education)+ 
  labs(
    title = "Total activity over the day by age",
    x = "Age",
    y = "Total activity",
    color = "Sex")

total_activity



#Total activity appears to decrease with age for individuals from all education levels. For those with less than high school or high school equivalent educational levels, males appear to have higher total activity than females at younger ages. Males and females with more than high school education levels have similar amounts of total activity at all ages but females have slightly more total activity at younger ages. 
```





Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences

```{r}

day_activity<-
  nhanes_long %>% 
  mutate(MIM_values=as.numeric(MIM_values))%>% 
  mutate(minute=as.numeric(minute))%>% 
  ggplot(aes(x=minute, y=MIM_values, color=sex))+
  geom_point(alpha=0.3)+
  geom_smooth(method = "lm")+
  facet_grid(education~.)+
  theme(strip.text.y = element_text(angle = 0))+ 
  labs(
    title = "MIM values over 24-hour period ",
    x = "Time (minutes)",
    y = "MIM values",
    color = "Sex")

#smooth trends were difficult to see alongside the points, creating separate smooth graphs 

day_activity_smooth<-
  nhanes_long %>% 
  mutate(MIM_values=as.numeric(MIM_values))%>% 
  mutate(minute=as.numeric(minute))%>% 
  ggplot(aes(x=minute, y=MIM_values, color=sex))+
  geom_smooth(method = "lm")+
  facet_grid(education~.)+
  theme(strip.text.y = element_text(angle = 0))+ 
  labs(
    title = "Linear model of MIM values over 24-hour period ",
    x = "Time (minutes)",
    y = "MIM values",
    color = "Sex")

day_activity/day_activity_smooth

#There is an overall increase for MIM values over a 24-hour period for all education levels and sexes. For males that have more than high school education level, there is a large peak in activity around the 1100-1300 minute range. There is a peak in activity for both sexes of the more than high school education level around the 500 minute mark. 

```








